---
title: "COVID-19 Cases in Ontario"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ONCOVID19 GitHub archive

The code in this archive uses the latest [public data on COVID-19 cases in Ontario](https://www.ontario.ca/page/2019-novel-coronavirus#section-0)

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Status of cases in Ontario

Ontario is now providing a more detailed daily epidemiologic summary of COVID-19 cases in the province ([January 15 - April 1, 2020](https://files.ontario.ca/moh-covid-19-report-en-2020-04-01.pdf)) that will be updated each day at 10:30 a.m. The new summary provides more provincial and regional data on confirmed cases, demographics and trends of cases since the outbreak began including geography, exposure and severity. 

## Confirmed positive cases of COVID19 in Ontario

```{r data}
url <- paste0("https://data.ontario.ca/dataset/",
              "f4112442-bdc8-45d2-be3c-12efae72fb27/resource/",
              "455fd63b-603d-4608-8216-7d8647f43350/download/",
              "conposcovidloc.csv")

dat_conf <- read.csv(url)

head(dat_conf)
```
